\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-04-25 mSystems/mSys_MAIN.tex
mSystems release 0.0.4
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/finalPlot33_new.pdf
Changes to mSystems Results section (rel0.0.5)
On branch master
	modified:   ../.gitignore
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
Daniel M. Martinez -> Daniel Martínez-Martínez
In mSystems move section from MandM to Results (v0.0.6)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.t

% release codenumber (update it here!)
\newcommand\releasecode{0.0.3}

@@

% text layout
\usepackage[letterpaper]{geometry}
\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space

% watermarks (time consuming!)
%\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
%\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
%\emph{mSystems} \LaTeX\ manuscript release \releasecode}

@@

% Importance environment and keywords command
\newcommand\importancename{Importance}
\newcommand{\importance}[1]{
    \small
    \begin{center}
        {\bfseries \importancename\vspace{-.5em}\vspace{0pt}}
    \end{center}
    \begin{quote}
    #1
    \end{quote}
}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

@@

\title{
	\vspace*{0mm} % Helps to include the corresponding author footnote in the title page
	\singlespacing
	\begin{flushleft}
		\texttt{\large Title:} \\
	\end{flushleft}
	\vspace*{2mm}
	Health and disease imprinted in the time variability of the human microbiome\\
	\vspace*{6mm}
	\begin{flushleft}
		\texttt{\large Running title:} \\
	\end{flushleft}
	\vspace*{0mm}
	Microbiota, are you sick?
	\vspace*{4mm}
	}

@@

\author[1,2]{Jose Manuel Martí}
\author[1,2,3]{Daniel M. Martínez}
\author[2]{Manuel Peña}
\author[1,2]{César Gracia}
\author[1,3,4,5]{Amparo Latorre}
\author[1,3,4,5]{Andrés Moya}
\author[1,2,\#]{Carlos P. Garay}

\affil[1]{Institute for Integrative Systems Biology (I2SysBio), 46980, Spain.}
\affil[2]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
\affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}
\affil[5]{CIBER en Epidemiología y Salud Pública (CIBEResp), Madrid, Spain}

@@

%TC:newcounter abswords Words in abstract
%TC:macro \abstract [abswords]
\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role.}

%TC:newcounter impwords Words in importance section
%TC:macro \importance [impwords]
\importance{Human microbiota is tightly associated to the health status of a person. Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months. Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas. This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system. }

\vspace{4mm}
\begin{keywords}
microbiome, systems biology, ecological modelling, community composition, stability	
\end{keywords}

>>>>

% release codenumber (update it here!)
\newcommand\releasecode{0.0.7}

@@

% text layout
%\usepackage[letterpaper]{geometry}
%\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
%\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space

% watermarks (time consuming!)
\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
\emph{mSystems} \LaTeX\ manuscript release \releasecode}

@@

% Importance environment and keywords command
\newcommand\importancename{Importance}
\newcommand{\importance}[1]{
	{
    \small
    \begin{center}
        {\bfseries \importancename\vspace{-.5em}\vspace{0pt}}
    \end{center}
    \begin{quote}
    #1
    \end{quote}
    }
}
\providecommand{\keywords}[1]{\footnotesize\textbf{\textit{Keywords---}} #1}

@@

\title{
	\vspace*{-16mm} % Helps to include the corresponding author footnote in the title page
	\singlespacing
	\begin{flushleft}
		\texttt{\large Title:} \\
	\end{flushleft}
	\vspace*{2mm}
	Health and disease imprinted in the time variability of the human microbiome\\
	\vspace*{6mm}
	\begin{flushleft}
		\texttt{\large Running title:} \\
	\end{flushleft}
	\vspace*{0mm}
	Microbiota, are you sick?
	\vspace*{4mm}
	}

@@

\author[1,2]{Jose Manuel Martí}
\author[1,2,3]{Daniel Martínez-Martínez}
\author[2]{Manuel Peña}
\author[1,2]{César Gracia}
\author[1,3,4,5]{Amparo Latorre}
\author[1,3,4,5]{Andrés Moya}
\author[1,2,\#]{Carlos P. Garay}

\affil[1]{Institute for Integrative Systems Biology (I2SysBio), 46980, Spain.}
\affil[2]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
\affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, UVEG, 46980, Spain.}
\affil[5]{CIBER en Epidemiología y Salud Pública (CIBEResp), Madrid, Spain}

@@

%TC:newcounter abswords Words in abstract
%TC:macro \abstract [abswords]
{\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role.}}

%TC:newcounter impwords Words in importance section
%TC:macro \importance [impwords]
\importance{Human microbiota is tightly associated to the health status of a person. Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months. Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas. This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.}

\vspace{4mm}
\begin{keywords}
microbiome, systems biology, ecological modeling, metagenomics, stability	
\end{keywords}

### 2016-04-25 mSystems/mSys_MandM.tex
Some work in mSys MAIN (fix font size) and Results (v0.0.7)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex
	modified:   results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_Rank.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_TimesCorr.pdf

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Section \ref{sec:X-w}). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{fig:taxlev2}
\end{figure}

>>>>

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{fig:taxlev2}
\end{figure}

### 2016-04-25 mSystems/mSys_results.tex
Some work in mSys MAIN (fix font size) and Results (v0.0.7)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex
	modified:   results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_Rank.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_TimesCorr.pdf

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time. We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
	\label{fig:main1}
\end{figure}

@@

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
	    \hline
	    \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}. This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10 and \bar{\beta} = 0.863 \pm 0.028$.}
  \label{tab:DK}
\end{table}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
	\caption{Microbiota states can be placed in the phase space $F$--$V$. The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75). We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability. Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days). We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes. We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
	\label{fig:main3}
\end{figure}

@@

%%%
\subsection*{DRAFT: Specific results}

\subsubsection*{Fit Plots}
Both for the unweighted fit (see Material and Methods section for details) and for the X-weighted fit (detailed in Material and Methods subsection), two plots are generated: the former with logarithmic scale, the latter with lineal scale. Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots. Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}. Among other tests, it allows to check for normality and homoscedasticity of the residues.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
	\label{fig:unwFit}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
	\label{fig:X-wFit}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}. The top-left subplot is a simple residues plot. The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided). The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot. Finally, the bottom-right subplot is a residues Normal histogram plot. This set of subplots allows to check for normality and homoscedasticity of the residues.}
	\label{fig:unwRes}
\end{figure}

\subsubsection*{Histogram Plots} 
\CC\ generates three different histogram plots:
\begin{description}
	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem. As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}. 
	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}. Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found. Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot. The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF. In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code). Figure \ref{fig:histZRF} shows an example of this plot. There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once. So, this plot is clearly useful to notice how the ``core'' is distributed.
\end{description}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
	\label{fig:histAFP}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:hist2D}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:histZRF}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
	\label{fig:corrElm}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
	\label{fig:corrTim}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:corrank}
\end{figure}

>>>>

\begin{figure}
	\centering
	\vspace*{-15mm} % Corrects overbox of the figures
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time. We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
	\label{fig:main1}
\end{figure}

@@

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
	    \hline
	    \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}. This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10$ and $\bar{\beta} = 0.863 \pm 0.028$.}
  \label{tab:DK}
\end{table}

@@

\begin{figure}
	\centering
	\vspace*{-5mm} % Corrects overbox of the figures
	\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
	\caption{Microbiota states can be placed in the phase space $F$--$V$. The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75). We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability. Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days). We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes. We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
	\label{fig:main3}
\end{figure}

@@

%%%
\subsection*{Specific results}

\subsubsection*{Fit Plots}
For each and every dataset included in the study, an unweighted fit (see Material and Methods section for details) and a X-weighted fit (detailed in Material and Methods subsection) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time. Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits. Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}). Among other tests, it allows to check for normality and homoscedasticity of the residues.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}}
	\label{fig:fits}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study. The top-left subplot is a simple residues plot. The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided). The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot. Finally, the bottom-right subplot is a residues Normal histogram plot. This set of subplots allows to check for normality and homoscedasticity of the residues.}
	\label{fig:unwRes}
\end{figure}

\subsubsection*{Histogram Plots (DRAFT)} 
\CC\ generates three different histogram plots:
\begin{description}
	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem. As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}. 
	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}. Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found. Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot. The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF. In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code). Figure \ref{fig:histZRF} shows an example of this plot. There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once. So, this plot is clearly useful to notice how the ``core'' is distributed.
\end{description}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
	\caption{2D histogram deviation plot of the data for patient A of the IBS study}
	\label{fig:hist2D}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study}
	\label{fig:histZRF}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf}
	\caption{Element correlation plot of the for the most dominant taxa in the data for patient A of the IBS study}
	\label{fig:corrElm}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods) in the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:corrank}
\end{figure}
